---
title: "Sandusky Bay 2024 Chlorophyll-a analysis"
author: "Ryan Wagner"
date: "`r Sys.Date()`"
output: "html_document"
editor_options: 
  chunk_output_type: console
---
<button class="back-button">Home</button>

<script>
    // Get references to the back button
    var backButton = document.querySelector('.back-button');
    // Define the function to navigate back to the index page
    const goBackToIndex = () => {
        window.location.href = "https://ryansw22.github.io/github.io/"; 
        // Replace "index.html" with the actual name of your index page
    };
    // Add event listener to the back button
    backButton.addEventListener('click', goBackToIndex);
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 5,
	fig.path = "./Figures/",
	fig.width = 7,
	message = FALSE,
	warning = FALSE,
	dpi = 300
)
rm(list=ls())
setwd("C:/Users/ryansw/Desktop/RSW files/2. Sandusky Bay/2024/Chlorophyll")
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# output: "github_document"
packages<-c("gplots", "readxl", "egg", "scales", "grid", "paletteer","ggtext",
            "reshape2", "RColorBrewer", "patchwork", "tidyverse")

installed_packages <- packages %in% rownames(installed.packages())
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]

if(length(new.packages)){
  install.packages(new.packages)
}
# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
rm("packages","installed_packages","new.packages")
theme_set(theme_article())
```

```{r Read in data, include=FALSE}
# Read in data.
SBCHL <- read_xlsx("./SB_2024_Chlorophyll-a.xlsx", sheet = "Raw")
# Convert date to actual date in r.
SBCHL$sample_date <- as.Date(SBCHL$sample_date,"%m/%d/%Y", tz = "GMT")
# Add in Year and DOY columns based on the sample date.
SBCHL<-
  SBCHL %>% 
  mutate(Year = year(sample_date),
         DOY = as.character(yday(sample_date)))
# Remove the "_" from the site names for a space.
SBCHL$site<- gsub("_"," ",SBCHL$site)
# Factor sites so they are in sampling order.
SBCHL$site<- factor(SBCHL$site, 
                    levels = c("Muddy Creek", "ODNR 4", "ODNR 6", "Bridge", "ODNR 2",
                               "Buoy 2", "ODNR 1", "EC 1163", "Causeway", "Bells",
                               "Causeway inside", "Causeway outside"))
# Create a y-axis label to call in plots.
name=expression("Chlorophyll-a Concentration"~("\u00B5"*g~L^"-1"))
```

::: {align="center"}
# Site time series
:::

```{r 2024_Chlorophyll, echo=FALSE, fig.align='center', results='asis'}
lab<-expression("Concentration"~("\U00B5"*g~L^"-1"))
cat("\n")
SBCHL %>%
      ggplot(aes(x=DOY,y= final_chl))+
        stat_summary(aes(x=DOY,y=final_chl, group = site, color = site),fun = "mean",geom = "point")+
        stat_summary(aes(x=DOY,y=final_chl, group = site, color = site),fun = "mean",geom = "line", show.legend = FALSE)+
        theme(axis.text.x = element_text(angle = 0,hjust = 0.5,vjust = 0.5, color = "black"),
              legend.position = "bottom",
              legend.title = element_blank(),
              axis.title.y = element_text(),
              plot.title = element_text(hjust = 0.5))+
        scale_y_continuous(labels = scales::number_format(accuracy = 1), limits = c(0,300),
                       breaks = seq(0, 300, length.out = 8), expand = expansion(c(0,0.01)))+
        ggtitle(paste0(unique(SBCHL$Year)," Sandusky Bay Chlorophyll-a"))+
        xlab(paste0("DOY ", "(", unique(SBCHL$Year),")"))+
        ylab(name)+
        scale_color_brewer(palette = "Paired")
```

::: {align="center"}
# Box plots by date
:::

#  {.tabset}

```{r Chlorophyll_day, echo=FALSE, fig.align='center', fig.path="./Figures/boxplot/", results='asis'}
for (i in unique(SBCHL$sample_date)) {
tmp <- SBCHL %>%
  filter(sample_date == i)
cat("## ",unique(format(tmp$sample_date, "%b %d, %Y")) , "\n")
tmpD<-unique(format(tmp$sample_date, "%b %d, %Y"))
  ptmp<-
  tmp %>% 
    ggplot(aes(x=site,y=final_chl))+
    geom_boxplot(aes(group=site))+
    scale_y_continuous(labels = scales::number_format(accuracy = 1), limits = c(0,300),
                       breaks = seq(0, 300, length.out = 8), expand = expansion(c(0,0.01)))+
    theme(axis.text.x = element_text(angle = 0,hjust = 0.5,vjust = 0.5, color = "black"),
          legend.title = element_blank(),
          plot.background = element_rect(fill = "white"),
          plot.title = element_text(hjust = 0.5))+
    labs(title = tmpD)+
    xlab("Site")+
    ylab(name)
cat("\n")
print(ptmp)
cat("\n")
cat("\n")
}
```

::: {align="center"}
# Bar plots by sample_date
:::

#  {.tabset}

```{r BPD, echo=FALSE, fig.align='center', fig.path="./Figures/barplot/", results='asis'}
for (i in unique(SBCHL$sample_date)) {
tmp <- SBCHL %>%
       filter(sample_date == i) 
cat("## ",unique(format(tmp$sample_date, "%b %d, %Y")) , "\n")
tmpD<-unique(format(tmp$sample_date, "%b %d, %Y"))
  tmp1<-
  tmp %>% 
    ggplot(aes(x=site))+
      stat_summary(aes(x=site,y=final_chl),fun = "mean",geom = "bar")+
      stat_summary(aes(x=site,y=final_chl),fun.data = "mean_sdl", 
            fun.args = list(mult = 1), width = 0.1, geom = "errorbar")+
      theme(axis.text.x = element_text(angle = 0,hjust=0.5,vjust=0.5, colour = "black"),
            legend.position = "none",
            legend.title = element_blank(),
            plot.background = element_rect(fill = "white"),
            plot.title = element_text(hjust = 0.5))+
      scale_y_continuous(breaks = scales::pretty_breaks(n=10),expand = expansion(c(0,0.01)))+
      labs(title = tmpD)+
      xlab("Site")+
      ylab(name)
cat("\n")
print(tmp1)
cat("\n")
cat("\n")
}
```