---
title: "Sandusky Bay 2024 Environmental analysis"
author: "Ryan Wagner"
date: "`r Sys.Date()`"
output: "html_document"
editor_options: 
  chunk_output_type: console
---
<button class="back-button">Home</button>

<script>
    // Get references to the back button
    var backButton = document.querySelector('.back-button');
    // Define the function to navigate back to the index page
    const goBackToIndex = () => {
        window.location.href = "https://ryansw22.github.io/github.io/"; 
        // Replace "index.html" with the actual name of your index page
    };
    // Add event listener to the back button
    backButton.addEventListener('click', goBackToIndex);
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	fig.height = 5,
	fig.path = "./Figures/",
	fig.width = 7,
	message = FALSE,
	warning = FALSE,
	dpi = 300
)
rm(list=ls())
setwd("C:/Users/ryansw/Desktop/RSW files/2. Sandusky Bay/2024/Environment")
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
packages<-c("readxl","egg","tidyverse","lubridate")

installed_packages <- packages %in% rownames(installed.packages())

new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]

if(length(new.packages)){
  install.packages(new.packages)
}
# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
rm("packages","installed_packages","new.packages")
theme_set(theme_article())
```

```{r Read in data, include=FALSE}
# Read in data
SBFIELD1 <- read_xlsx("./SB_2024_Environmental.xlsx", sheet = "Condensed")
# Convert date to actual date in r.
SBFIELD1$Date <- as.Date(SBFIELD1$Date,"%m/%d/%Y")
# Convert non-numeric values that are supposed to be, to numeric.
SBFIELD1 <- SBFIELD1 %>% mutate_at(3:17,as.numeric) %>% mutate_at("Secchi", as.numeric)
# Replaces the "_" with a space for all Sites.
SBFIELD1$Site<- gsub("_"," ",SBFIELD1$Site)
# Factors the sites so they will appear in the correct sample order.
SBFIELD1$Site<- 
  factor(SBFIELD1$Site, levels = c("Muddy Creek","ODNR 4","ODNR 6","Bridge","ODNR 2","Buoy 2","ODNR 1","EC 1163",
                                  "Causeway","Bells","Inside Buoy","Outside Buoy"))
SBFIELD<-
SBFIELD1 %>%
  mutate(Year = year(Date),# Create a column for year.
         DOY = as.character(yday(Date))) %>% # Create a column for DOY.
  relocate(DOY, .before = "Air_Temperature") %>% 
  relocate(Year, .after = "DOY") %>% 
  pivot_longer(5:19, names_to = "variable",values_to = "count")# Pivots data from a wide format to long to make plotting easier later.
SBFIELD$variable<- gsub("_"," ",SBFIELD$variable)# Removes the "_" for a space in the newly created variable column

```

::: {align='center'}
# Environmental Parameters
:::

# {.tabset}

```{r Airtemp, echo=FALSE,results='asis'}
lab<-expression("Air Temperature"~("\U00B0"*C))
tmp<-SBFIELD %>% 
 filter(variable=="Air Temperature") %>%  
  filter(!grepl("sed", Site)) %>% 
  na.omit()
cat("## ", unique(tmp$variable), "\n")  
Air.Temp.plot<-tmp %>% 
    ggplot()+
      geom_point(aes(x = DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            plot.caption = element_text(hjust = 0),
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab(lab)+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 1), limits = c(13,29),
                          breaks = seq(13, 29, length.out = 8))
cat("\n")
print(Air.Temp.plot)
cat("\n")
cat("\n")
```

```{r secchi, echo=FALSE,results='asis'}
lab<-expression("Secchi"~("cm"))
tmp<-SBFIELD %>% 
 filter(variable=="Secchi") %>% 
  na.omit()
cat("## ", unique(tmp$variable), "\n")  
Secchi.plot<-tmp %>% 
    ggplot()+
      geom_point(aes(x = DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            legend.title.align = 0.5,
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab(lab)+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 0.5), limits = c(10, 120),
                          breaks = seq(10, 120, length.out = 8))
cat("\n")
print(Secchi.plot)
cat("\n")
cat("\n")
```

```{r Watertemp, echo=FALSE,results='asis'}
lab<-expression("Water Temperature"~("\U00B0"*C))
tmp<-SBFIELD %>% 
  filter(variable=="Water Temperature") %>%  
  na.omit()
cat("## ", unique(tmp$variable), "\n")
Water.Temp.plot<- tmp %>% 
    ggplot()+
      geom_point(aes(x =DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            legend.title.align = 0.5,
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab(lab)+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 0.5), limits = c(17,27),
                          breaks = seq(17,27, length.out = 8))
cat("\n")
print(Water.Temp.plot)
cat("\n")
cat("\n")
```

```{r DO, echo=FALSE,results='asis'}
lab<-expression("Dissolved Oxygen"~("mg"~L^"-1"))
tmp<-SBFIELD %>% 
  filter(variable=="Dissolved Oxygen") %>%
  na.omit()
cat("## ", unique(tmp$variable), "\n")
DO.plot<-tmp %>% 
    ggplot()+
      geom_point(aes(x = DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            legend.title.align = 0.5,
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab(lab)+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 0.5), limits = c(4.5, 13),
                          breaks = seq(4.5, 13, length.out = 8))
cat("\n")
print(DO.plot)
cat("\n")
cat("\n")
```

```{r cond, echo=FALSE,results='asis'}
lab<-expression("Conductivity"~("\U00B5"*S~cm^"-1"))
tmp<-SBFIELD %>% 
  filter(variable=="Conductivity") %>% 
  na.omit()
cat("## ", unique(tmp$variable), "\n")  
Cond.plot<-tmp %>% 
    ggplot()+
      geom_point(aes(x = DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            legend.title.align = 0.5,
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab(lab)+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 1), limits = c(200, 700),
                          breaks = seq(200, 700, length.out = 8))
cat("\n")
print(Cond.plot)
cat("\n")
cat("\n")
```

```{r pH, echo=FALSE,results='asis'}
tmp<-SBFIELD %>% 
  filter(variable=="pH") %>%
  na.omit()
cat("## ", unique(tmp$variable), "\n")
pH.plot<-tmp %>% 
    ggplot()+
      geom_point(aes(x = DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            legend.title.align = 0.5,
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab("pH")+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 0.05), limits = c(7.5,9.5),
                          breaks = seq(7.5,9.5, length.out = 8))
cat("\n")
print(pH.plot)
cat("\n")
cat("\n")
```

```{r Turbidity, echo=FALSE,results='asis'}
lab<-expression("Turbidity"~("NTU"))
tmp<-SBFIELD %>% 
 filter(variable=="Turbidity") %>% 
  na.omit()
cat("## ", unique(tmp$variable), "\n")  
Turb.plot<-tmp %>% 
    ggplot()+
      geom_point(aes(x = DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab(lab)+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 1), limits = c(0, 350),
                          breaks = seq(0, 350, length.out = 8))
cat("\n")
print(Turb.plot)
cat("\n")
cat("\n")
```

```{r PC, echo=FALSE,results='asis'}
lab<-expression("Phycocyanin"~("\u00B5"*g~L^"-1"))
tmp<-SBFIELD %>% 
 filter(variable=="Phycocyanin Sonde") %>% 
  filter(!grepl("sed", Site)) %>%  
  na.omit()
cat("## ", unique(tmp$variable), "\n")  
PC.plot<-tmp %>% 
    ggplot()+
      geom_point(aes(x = DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab(lab)+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 0.5), limits = c(0, 8),
                          breaks = seq(0, 8, length.out = 8))
cat("\n")
print(PC.plot)
cat("\n")
cat("\n")
```

```{r CHL, echo=FALSE,results='asis'}
lab<-expression("Chlorophyll-a Concentration"~("\u00B5"*g~L^"-1"))
tmp<-SBFIELD %>% 
 filter(variable=="Chlorophyll Sonde") %>% 
  na.omit()
cat("## ", unique(tmp$variable), "\n")  
CHL.plot<-tmp %>% 
    ggplot()+
      geom_point(aes(x = DOY, y = count, group = Site, color = Site), size = 2 )+
      geom_line(aes(x = DOY, y = count, group = Site, color = Site),linewidth=1,show.legend = FALSE)+
      theme(axis.text.x=element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
            legend.position = "right",
            plot.caption = element_text(hjust = 0),
            plot.background = element_rect(fill = "white"))+
      scale_color_brewer(palette = "Paired")+
      ylab(lab)+
      labs(caption = "Measured with Sonde")+
      xlab(paste0("DOY ","(",unique(tmp$Year),")"))+
      scale_y_continuous(labels = scales::number_format(accuracy = 1), limits = c(0,90),
                          breaks = seq(0, 90, length.out = 8))
cat("\n")
print(CHL.plot)
cat("\n")
cat("\n")
```


