---
title: "Sandusky Bay Light 2024 Light Attenuation analysis"
author: "Ryan Wagner"
date: "`r Sys.Date()`"
output: "html_document"
editor_options: 
  chunk_output_type: console
---
<button class="back-button">Home</button>

<script>
    // Get references to the back button
    var backButton = document.querySelector('.back-button');
    // Define the function to navigate back to the index page
    const goBackToIndex = () => {
        window.location.href = "https://ryansw22.github.io/github.io/"; 
        // Replace "index.html" with the actual name of your index page
    };
    // Add event listener to the back button
    backButton.addEventListener('click', goBackToIndex);
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 5,
	fig.path = "./Figures/",
	fig.width = 7,
	message = FALSE,
	warning = FALSE,
	dpi = 300
)
rm(list = ls())
setwd("C:/Users/ryansw/Desktop/RSW files/2. Sandusky Bay/2024/PAR Light")
```

```{r warning=FALSE, include=FALSE}
packages<-c("readxl", "egg", "tidyverse","RColorBrewer","scales",
            "patchwork","lubridate","reshape2","rLakeAnalyzer","rmarkdown")
for(i in packages){
  if(i %in% rownames(installed.packages()) == FALSE){
    install.packages(i)
    if(i %in% rownames(installed.packages()) == TRUE){
    }
  }
  invisible(lapply(i, library, character.only = TRUE))
}
theme_set(theme_article())
rm("i","packages")
```

```{r include=FALSE}

LIGHT<-read_excel("SB_2024_PAR Light.xlsx", sheet = "Light")
# convert columns to numeric, pivot to a longer df for plotting, and pull light value at surface into Light_0 column for calcs later.
LIGHT<-
  LIGHT %>% 
    filter(Rep=="avg") %>% 
    group_by(Site,Date) %>%
    select(-c("Notes","Rep")) %>% 
    mutate_at(4:15, as.numeric) %>% 
    pivot_longer(cols = 4:15, names_to = "Depth",values_to = "LightValue") %>% 
    drop_na()  %>%
    mutate_at("Depth",as.numeric) %>% 
    mutate_at("LightValue",as.numeric) %>% 
    mutate_at("Date", as.Date) %>% 
    filter(LightValue>0) %>% 
    mutate(Light_0=case_when(Depth==0~LightValue,TRUE~NA)) %>%
    fill(Light_0) %>%
    ungroup()
# Remove the underscore from site names.
LIGHT$Site<-gsub("_"," ",LIGHT$Site)
# Organize the sites in order of sampling.
LIGHT$Site<-factor(LIGHT$Site, 
                   levels = c("Muddy Creek", "ODNR 4", "ODNR 6", "ODNR 2", "Buoy 2", 
                              "ODNR 1", "EC 1163",  "Bells", "Causeway"))
# Create a label for the y- axis to call in plots
lab=expression("Photosynthetic Photon Flux Density (PPFD)"~"("*"\u00B5"*mol~m^-2~sec^-1*")")
# Light attenuation equation: kd = 1/z ln(E0 /Ez ) 
# kd is the light coefficient between the surface and z depth
L2<- 
  LIGHT %>% 
    mutate(Year=year(Date),.before = DOY) %>% 
    mutate(Month=month(Date),.before = Year) %>% 
    mutate(kd=(-1/Depth)*log(LightValue/Light_0)) %>% 
    na.omit() %>% 
# Group by site and DOY to get daily Kd
    group_by(Site, DOY) %>% 
# K_daily_avg is the light coefficient average for all depths
    mutate(K_daily_avg= mean(kd)) %>% 
    ungroup() %>%
# Group by year and site to get yearly k values
    group_by(Site,Year) %>% 
    mutate(K_avg=mean(K_daily_avg)) %>% 
    ungroup() 

```


::: {align='center'}
# Daily light extinction coefficients
:::

```{r Yearly_K, results='asis'}
T1<-L2 %>% select(-c(Date,Light_0, kd,K_avg,Year,Depth,LightValue)) %>% 
  group_by(Site,DOY) %>% 
  slice(1:1) %>% 
  arrange(Site)
# This for loop will give individual site k values for the year
# for (i in unique(T1$site)) {
#   tmp<- T1 %>% filter(site== i) %>% mutate_at("DOY", as.character)
#   
#   kp1<-
#   tmp %>% 
#       ggplot() +
#         geom_point(aes(x=DOY, y = K_daily_avg), stat = "identity", size = 2)+
#         geom_line(aes(x=DOY, y = K_daily_avg), group = DOY, linewidth = 1.05)+
#         scale_y_continuous(labels = scales::number_format(accuracy = 0.5), 
#                                limits = c(0,max(tmp$K_daily_avg)),
#                                expand = expansion(c(0,0.01)))+
#         ggtitle(i)+
#         xlab("Day of Year (2023)")+
#         ylab(bquote(k[d]~(m^-1)))
# print(kp1)
# }
  T1 %>% 
    group_by("Site") %>% 
    mutate_at("DOY", as.character) %>% 
      ggplot() +
        geom_point(aes(x=DOY, y = K_daily_avg, color = Site), stat = "identity", size = 2)+
        geom_line(aes(x=DOY, y = K_daily_avg,group = Site, color = Site), linewidth = 1.05, show.legend = F)+
        scale_y_continuous(labels = scales::number_format(accuracy = 0.5), 
                           limits = c(0,max(T1$K_daily_avg)),
                           breaks = seq(0,max(T1$K_daily_avg), length.out = 8),
                           expand = expansion(c(0,0.01)))+
        ggtitle("Sandusky Bay light attenuation coefficients")+
        xlab("Day of Year (2024)")+
        ylab(bquote(k[d]~(m^-1)))+
        theme(axis.title.y = element_text(size = 12),
              legend.title.align = 0.5)


```

::: {align='center'}
# Daily Light attenuation curves for all sites
:::
# {.tabset}

```{r DLA, echo=FALSE, fig.path="./Figures/Daily Light Attenuation/" , results='asis'}
all_list<-list()
for (d in unique(L2$Date)) {
  L3<-L2 %>% filter(Date==d)
  DOY <- unique(L3$DOY)
  my_list<-list()
cat("## ", unique(L3$DOY), "\n", "\n")
for (i in unique(L3$Site)) {
  tmp<-
      L3 %>% 
        filter(Site==i) %>% 
        group_by(Depth) %>% 
        slice(1:1) %>% 
        select(-c(Site,Light_0,kd))%>% 
        filter(LightValue>0) %>% 
        mutate(Site=i) %>% 
        ungroup()
  # calculate 1% light depth
    ## first fit a liner model to the logarithmic light data
    m <- lm(log(tmp$LightValue) ~ tmp$Depth)
    ## then calculate the depth where 1% light is left
    z_light <- log(0.01) / m$coefficients[2]
    eqt <- paste0("y = ", round(coef(m)[1], 2), " ",
           round(coef(m)[2], 2), "x")
    ## Create a text grob for adding linear model equation to plot
    grob <- grid::grobTree(grid::textGrob(eqt, x = 0.1,  y= 0.95, hjust= 0, 
                              gp = grid::gpar(col="red", fontsize=13, fontface="bold")))
    ## Adds a column to tmp for the 1%light value and adds a variable column with light in it to get a legend.
    tmp<- tmp %>% mutate(z_light = z_light) %>% mutate(variable = "Light")

  # create the light plots
  # add data to list for saving
    my_list[[i]]<-as.data.frame(tmp)
    
  p1 <- tmp %>%  
          ggplot(aes(x = Depth, y = LightValue, color = variable)) +
            geom_line(linewidth=1.05) + 
            geom_point(size=2) + 
            coord_flip() +
            theme(legend.position = "bottom",
                  legend.title = element_blank()) + 
            xlab("Depth (m)")  +
            scale_x_continuous(labels = scales::number_format(accuracy = 0.5),
                               trans = "reverse",
                               breaks = seq(min(tmp$Depth), max(tmp$Depth), by = 0.5))+
            ggtitle(paste(i,"Light Attenuation", format(as.Date(d,origin),"%m/%d/%Y"))) +
            ylab(lab)+
            scale_y_continuous(labels = scales::number_format(accuracy = 1), 
                               limits = c(0,max(tmp$LightValue)),
                               breaks = seq(0,max(tmp$LightValue), length.out = 8),
                               expand = expansion(c(0,0.01)))+
            geom_vline(data = data.frame(x = z_light, variable = "1% light depth"),
                       group = "Lightvalue",
                       aes(xintercept = z_light, color = variable ), linetype = "dashed")
    cat("\n")
    print(p1)
    cat("\n")
    cat("\n")

}
all_list[[as.character(DOY)]]<-my_list
}
```

::: {align='center'}
# Individual sites yearly light curves
:::
# {.tabset}

```{r Light_Curve, echo=FALSE, fig.align='center', fig.path="./Figures/Yearly Light Attenuation/" ,  results='asis'}

for (i in unique(L2$Site)) {
  tmp<-
      L2 %>% 
        filter(Site==i) %>% 
        group_by(Depth) %>% 
        mutate(LightValue=mean(LightValue)) %>% 
        slice(1:1) %>%
        filter(LightValue>0) %>%
        select(-c(Site,Light_0,kd))%>% 
        mutate(Site=i) %>% 
        ungroup()
  # calculate 1% light depth
    ## first fit a liner model to the logarithmic light data
    m <- lm(log(tmp$LightValue) ~ tmp$Depth)
    ## then calculate the depth where 1% light is left
    z_light <- log(0.01) / m$coefficients[2]
    eqt <- paste0("y = ", round(coef(m)[1], 2), " ",
           round(coef(m)[2], 2), "x")
    ## Create a text grob for adding linear model equation to plot
    grob <- grid::grobTree(grid::textGrob(eqt, x = 0.1,  y= 0.95, hjust= 0, 
                              gp = grid::gpar(col="red", fontsize=13, fontface="bold")))
    ## Adds a column to tmp for the 1%light value and adds a variable column with light in it to get a legend.
    tmp<- tmp %>% mutate(z_light = z_light) %>% mutate(variable = "Light")
    my_list[[i]]<-as.data.frame(tmp)
cat("## ", i, "\n","\n")   
  p1 <- tmp %>%  
          ggplot(aes(x = Depth, y = LightValue, color = variable)) +
            geom_line(linewidth=1.05) + 
            geom_point(size=2) + 
            coord_flip() +
            theme(legend.position = "bottom",
                  legend.title = element_blank()) + 
            xlab("Depth (m)")  +
            scale_x_continuous(labels = scales::number_format(accuracy = 0.5),
                               trans = "reverse",
                               breaks = seq(min(tmp$Depth), max(tmp$Depth), by = 0.5))+
            ggtitle(paste(i,"Light Attenuation")) +
            ylab(lab)+
            scale_y_continuous(labels = scales::number_format(accuracy = 1), 
                               limits = c(0,max(tmp$LightValue)),
                               breaks = seq(0,max(tmp$LightValue), length.out = 8),
                               expand = expansion(c(0,0.01)))+
            geom_vline(data = data.frame(x = z_light, variable = "1% light depth"),
                       group = "Lightvalue",
                       aes(xintercept = z_light, color = variable ), linetype = "dashed")
    cat("\n")
    print(p1)
    cat("\n")
    cat("\n")

}

```
